### 2020-02-27 drugspending.Rmd
medicare drug spending cleaned

---
title: "Medicare Part B Drug Spending"
author: "Suruchi Ramanujan"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(janitor)
```

```{r}
drugspending<-read.csv("Medicare Part B Drug Spending YTD 2017.csv", skip = 3) %>%
clean_names() %>%
  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units",
         ,"x2014_total_spending", "x2014_total_dosage_units", ,"x2015_total_spending", "x2015_total_dosage_units", ,"x2016_total_spending", "x2016_total_dosage_units", ,"x2017_total_spending", "x2017_total_dosage_units",
         "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
```