### 2020-02-28 drugspending.Rmd
drug spending and opioid death distribution

---
title: "Medicare Part B Drug Spending"
author: "Suruchi Ramanujan"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(janitor)
```

```{r}

# This data tracks the spending over a course of five years of a series of
# opioids. Perhaps, we will find a correlation between spending and drug type
# for opioid deaths.

drugspending<-read.csv("Medicare Part B Drug Spending YTD 2017.csv", skip = 3) %>%
clean_names() %>%
  select("drug_description", "brand_name", "generic_name","x2013_total_spending", "x2013_total_dosage_units", "x2014_total_spending", "x2014_total_dosage_units", "x2015_total_spending", "x2015_total_dosage_units",
  "x2016_total_spending", "x2016_total_dosage_units", "x2017_total_spending", "x2017_total_dosage_units",
         "change_in_average_spending_per_dosage_unit_2016_2017", "annual_growth_rate_in_average_spending_per_dosage_unit_2013_2017")
```

```{r}

# This data set displays the number of deaths per county in NY over a course of
# 15 years.

deathbycounty_ny<- read.csv("Vital_Statistics__Opioid-Related_Deaths_by_County__Beginning_2003.csv") %>%
  clean_names()
```

Introduction: This project is meant to look into trends regarding the issue of opioid overuse. The misuse of opioids has ignited a national crisis that has contributed to worsening health outcomes for thousands and their families throughout the United States. This

Data set 1:

Data set 2:

Data set 3:

Data source: https://about.twitter.com/en_us/values/elections-integrity.html#data

About Me: My name is Suruchi Ramanujan and I am a senior in Quincy House studying Molecular and Cellular Biology. I grew up in Overland Park, Kansas so this year has been a real win for me (Go Chiefs). In my free time, I enjoy performing Indian Classical dance, running, and grinding through my DataCamp exercises.

My github repo for this project is below: