### 2020-03-20 ms_5.Rmd
new ggplots!!

---
title: "ms_5"
author: "Suruchi Ramanujan"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(janitor)
library(tidyverse)
```

*Copy pasted from last week
```{r}

# This first data set is the new data set from last week. This data provides
# state-specific numbers from the CDC regarding the number of drug overdose
# deaths per state as well as the predicted numbers for drug overdose deaths. I
# am super excited to plot this graph onto a map and look at how the worst
# affected states change over time.

drugoverdosedeaths<- read.csv(
  "VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>%
  clean_names() %>%
  select("state", "year", "month", "indicator", 
         "data_value", "state_name", "predicted_value") %>%

# I only want the information rows containing drug death information- I still
# need to decipher the difference between "Deaths" and "Drug Overdose Deaths"
# though.

 filter(str_detect(indicator, "Death")) %>%

# Ultimately, I want to look at state-specific data/trends so I want to group by
# state.

 group_by("state_name")
```

*New for this week
```{r}
# For this week, I am going to plot the Drug Overdose Deaths in April of 2015 by
# state according to the CDC. I want to use this data just because it is the
# first recorded date of drug overdose deaths. I am hoping there will be some
# expected trend across states.

drugoverdosedeaths$data_value <-as.numeric(drugoverdosedeaths$data_value)

drugodapr15<- drugoverdosedeaths %>%
  arrange(desc(data_value)) %>%
  select(state_name, `"state_name"`, data_value, indicator, month, year) %>%
  filter(year == "2015",
        month == "April",
        indicator == "Number of Drug Overdose Deaths")

ggplot(drugodapr15, aes(x= reorder(state_name,-data_value), y= data_value)) +
geom_col() +
labs(title = "Drug Overdose Deaths by State (April 2015)", 
     x = "State", 
     y = "Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}

# I am going to do the same with July 2019, which is the latest date of
# recording.

drugodjul19<- drugoverdosedeaths %>%
  arrange(desc(data_value)) %>%
  select(state_name, `"state_name"`, data_value, indicator, month, year) %>%
  filter(year == "2019",
        month == "July",
        indicator == "Number of Drug Overdose Deaths")

ggplot(drugodjul19, aes(x= reorder(state_name,-data_value), y= data_value)) +
geom_col() +
labs(title = "Drug Overdose Deaths by State (July 2019)", 
     x = "State", 
     y = "Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
difference_over_time<- drugodapr15 %>%
  full_join(drugodjul19, by = "state_name") %>%
  mutate(difference = data_value.y - data_value.x)

ggplot(difference_over_time, aes(x= reorder(state_name,-difference), y= difference)) +
geom_col() +
labs(title = "Difference in Drug Overdose Deaths between April 2015 and July 2019", 
     x = "State", 
     y = "Difference in Drug Overdose Deaths") +
theme_classic()+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))
```