### 2020-04-26 ms_5.Rmd
just need to fix legends in app/titles

---
title: "ms_5"
author: "Suruchi Ramanujan"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}

# I want to load all necessary packages.

library(dplyr)
library(janitor)
library(tidyverse)
library(ggthemes)
library(maps)
library(leaflet)
library(gganimate)
library(broom)
```

```{r}

# total deaths per year in u.s. by cancer, drug overdose, cvd

deathsbycommondiseases <- read_csv("deathsbycommondiseases.csv", col_types = "ldddddd") %>%
  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
  mutate(`Proportion of 1999 Deaths by Cancer` = `Deaths by Cancer`/549829) %>%
  mutate(`Proportion of 1999 Deaths by Drugs` = `Deaths by Drug Overdose`/19122) %>%
  mutate(`Proportion of 1999 Deaths by CVD` = `Deaths by CVD`/949900)

totaldeathsbycommondiseases <- deathsbycommondiseases %>%
  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, `Deaths by CVD`) %>%
  pivot_longer(.,
             cols = starts_with("Deaths by"),
             names_prefix = "Deaths by",
             values_to = "Deaths") %>%
  rename(`Cause of Death` = name)

propdeathsbycommondiseases <- deathsbycommondiseases %>%
  select(Year, `Proportion of 1999 Deaths by Cancer`, 
         `Proportion of 1999 Deaths by Drugs`, 
         `Proportion of 1999 Deaths by CVD`) %>%
  pivot_longer(.,
             cols = starts_with("Proportion of"),
             names_prefix = "Proportion of 1999 Deaths by",
             values_to = "Prop.Deaths") %>%
  rename(`Cause of Death` = name)
```

```{r}
plot_deathsbycommondiseases <- ggplot(totaldeathsbycommondiseases, aes(x = Year, y = Deaths, color = `Cause of Death`)) +
  geom_line() +
  scale_color_viridis_d() +
  theme_classic() +
   labs(x = "Year",
       y = "Total Number of Deaths in the United States") +
  geom_point() +
  transition_reveal(Year)

anim_save("plot_deathsbycommondiseases.gif", plot_deathsbycommondiseases)

prop_deathsbycommondiseases <- ggplot(propdeathsbycommondiseases, aes(x = Year, y = Prop.Deaths, color = `Cause of Death`)) +
  geom_line() +
  scale_color_viridis_d() +
  theme_classic() +
  labs(x = "Year",
       y = "Proportion of 1999 Total Deaths in the United States") +
  geom_point() +
  transition_reveal(Year)

anim_save("prop_deathsbycommondiseases.gif", prop_deathsbycommondiseases)

```

```{r}

# https://wonder.cdc.gov/controller/datarequest/D77;jsessionid=3F8131C534663BDD3F79E66994FCF12E

death_by_year_and_state <- read_csv("Drug Overdose Deaths by Year and State.csv")

us_state <- map_data("state") %>%
  rename(state_full = region)

death_by_year_and_state <- death_by_year_and_state %>%
  filter(`UCD - Drug/Alcohol Induced Code` == "D") %>%
  rename(state_full = State) %>%
  mutate(state_full = tolower(state_full)) %>%
  left_join(us_state, by = "state_full" )

death_by_year_and_state$`Crude Rate` <-as.numeric(death_by_year_and_state$`Crude Rate`)

ggplot(data = death_by_year_and_state,
            mapping = aes(x = long, y = lat,
                          fill = `Crude Rate`, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A") +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

```

```{r, message=FALSE}
#drug treatment centers

treatment_locations <- read_csv("Drug Treatment Centers.csv") %>%
  select(`Program Name`, Street, City, State, Zipcode)

# state populations

state_pop <- read_csv("State Population Estimate.csv") %>%
  rename(state_full = NAME) %>%
  mutate(state_full = tolower(state_full))

# https://simplemaps.com/data/us-cities

longlatinfo <- read_csv("uscitieslonglat.csv") %>%
  rename(City = city) %>%
  rename(State = state_id)

treatment_locations_map <- treatment_locations %>%
  left_join(longlatinfo, by = c("City", "State")) %>%
  group_by(State) %>%
  mutate(count = n()) %>%
  mutate(state_name = tolower(state_name)) %>%
  rename(state_full = state_name) %>%
  left_join(us_state, by = "state_full")

# not in final

ggplot(data = treatment_locations_map,
            mapping = aes(x = long, y = lat.y,
                          fill = count, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,120)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom") +
  labs(title = "Distribution of Opioid Treatment Centers in 2020",
       color = "Number of Treatment Centers in State")

treatment_locations_map_per_cap <- treatment_locations_map %>%
  left_join(state_pop, by = "state_full") %>%
  group_by(State) %>%
  mutate(count_per_pop = n()/POPESTIMATE2019)

```

```{r}
# in final

ggplot(data = treatment_locations_map_per_cap,
            mapping = aes(x = long, y = lat.y,
                          fill = count_per_pop, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,120)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom") +
  labs(title = "Distribution of Opioid Treatment Centers in 2020 (Per Capita)",
       color = "Number of Treatment Centers in State Per Capita")

```

```{r, message = FALSE}
#MA specific data

madeathbyagegender <- read_csv("MAFatalOpioidOverdosesByAgeAndGender.csv")

macountyheroin <- read_csv("MAHeroinAsPrimarySubstanceOfUse.csv")

madeathbyrace <- read_csv("MAOpioidDeathsByRaceAndEthnicity.csv")

matypeofsubstance <- read_csv("MAPrimarySubstanceofUseWhenEnteringTreatmentbyTown.csv")

toxdeath <- read_csv("MAProportionDecedentsWithRxHistorybyToxicologyScreen.csv",
         skip = 1)

```

```{r}
mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv", col_types = "cdc")

mavsus_death <- mavsus_death %>%
  group_by(Geography) %>%
  filter (Year != "2015") %>%
  filter (Year != "1999") %>%
  rename(deathperhundredthousand =`Age-Adjusted Opioid-Related Death Rate per 100,000 People`)

plot_mavsus_death <- ggplot(mavsus_death, aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
  geom_line() +
  scale_color_viridis_d() +
theme_classic() +
  labs(x = "Year",
       y = "Deaths per 100,000",
       title = "Comparison of Death Rates between Massachusetts and the United States",
       subtitle = "Between the Years of 2000 and 2014") +
  geom_point() +
  transition_reveal(Year)

anim_save("plot_mavsus_death.gif", plot_mavsus_death)

```

```{r}
counties <- read_csv("zip_codes_states.csv") %>%
  filter(state == "MA") %>%
  rename(County = county) %>%
  rename(Municipality = city)

madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv")

countypop <- read_csv("countypop.csv") %>%
  mutate(Pop = Pop/100000) %>%
  rename(subregion = CTYNAME) %>%
  mutate(subregion = tolower(subregion)) %>%
  select(subregion, Pop)

madeathbycountywlonglat <- madeathbycounty %>%
  left_join(counties, by = "Municipality")

madeathbycountywlonglat <- madeathbycountywlonglat %>%
  select(County,
         `Confirmed Opioid Related Death Count 2001-2005`,
         `Confirmed Opioid Related Death Count 2006-2010`,
         `Confirmed Opioid Related Death Count 2011-2015`,
         `latitude`,
         `longitude`,
         `Municipality`) %>%
  distinct(Municipality, .keep_all= TRUE) %>%
  na.omit() %>%
  rename(subregion = County) %>%
  mutate(subregion = tolower(subregion)) %>%
  group_by(subregion) %>%
  mutate(total_deaths_2001.5 = sum(`Confirmed Opioid Related Death Count 2001-2005`)) %>%
  mutate(total_deaths_2006.10 = sum(`Confirmed Opioid Related Death Count 2006-2010`)) %>%
  mutate(total_deaths_2011.15 = sum(`Confirmed Opioid Related Death Count 2011-2015`)) %>%
  distinct(subregion, .keep_all= TRUE)

us_county <- map_data("county") %>%
  filter(region == "massachusetts")

# https://www.indexmundi.com/facts/united-states/quick-facts/massachusetts/percent-of-people-of-all-ages-in-poverty#table

povertybycounty <- read_csv("PovertyByCounty.csv") %>%
  rename(subregion = County) %>%
  mutate(subregion = tolower(subregion))

full_data <- us_county %>%
  left_join(madeathbycountywlonglat, by = "subregion") %>%
  left_join(countypop, by = "subregion") %>%
  left_join(povertybycounty, by = "subregion") %>%
  mutate(percap_2001.5 = total_deaths_2001.5/Pop) %>%
  mutate(percap_2006.10 = total_deaths_2006.10/Pop) %>%
  mutate(percap_2011.15 = total_deaths_2011.15/Pop)

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2001.5)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

full_data %>%
  summarize(correlation = cor(percap_2001.5,`Poverty Percent`))

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2006.10)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

full_data %>%
  summarize(correlation = cor(percap_2006.10,`Poverty Percent`))

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2011.15)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

full_data %>%
  summarize(correlation = cor(percap_2011.15,`Poverty Percent`))

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = `Poverty Percent`,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,2,4,6,8,10)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2001.5,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2006.10,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2011.15,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

# don't put these in final app

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2001.5,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2006.10,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2011.15,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

```

```{r}

#treatment centers in MA

treatment_locations_map_mas <- treatment_locations %>%
  left_join(longlatinfo, by = c("City", "State")) %>%
  filter(State == "MA") %>%
  drop_na()

leaflet(options = leafletOptions(dragging = TRUE,
                                 minZoom = 8, 
                                 maxZoom = 10)) %>%
  addProviderTiles("CartoDB") %>%
  addCircleMarkers(data = treatment_locations_map_mas,
                   radius = 3,
                   label = ~`Program Name`)

# need to figure this out bc color isnt working
 # addLegend(position = "bottomright", colors = "#0000FF", title = "Opioid Treatment Sites")

```

```{r}

# models generated from data from kff.org

# ageopioid

ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", col_types = "dcc") %>%
  clean_names() %>%
  na.omit() %>%
  filter(age_range != "Total") %>%
  mutate(age_range = as.factor(age_range)) %>%
    mutate(number_of_deaths = as.numeric(number_of_deaths))

raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", col_types = "dcc") %>%
  clean_names() %>%
  na.omit() %>%
  mutate(race = as.factor(race)) %>%
  mutate(opioid_deaths = as.numeric(opioid_deaths))

ageandracemodel <- ageopioidmodel %>%
  na.omit() %>%
  full_join(raceopioidmodel, by = "year") %>%
  rename(deathsbyage = number_of_deaths) %>%
  rename(deathsbyrace = opioid_deaths)

ggplot(ageandracemodel, aes(year, deathsbyage, color = age_range)) +
  geom_point() +
  scale_color_viridis_d() +
  theme_classic() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Age Range")

ggplot(ageandracemodel, aes(year, deathsbyrace, color = race)) +
  geom_point() +
  scale_color_viridis_d() +
  theme_classic() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Race")
```

```{r}
# filtering data by race to show effects of time on deaths in each race group

death_vs_white <- raceopioidmodel %>%
  filter(race == "White, Non-Hispanic")

death_vs_white %>%
    lm(opioid_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high) %>%
    rename(Coefficient = estimate) %>%
    rename(`Lower End` = conf.low) %>%
    rename(`Upper End` = conf.high) %>%
    gt()

death_vs_black <- raceopioidmodel %>%
  filter(race == "Black, Non-Hispanic")

death_vs_black %>%
    lm(opioid_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

death_vs_hispanic <- raceopioidmodel %>%
  filter(race == "Hispanic")

death_vs_hispanic %>%
    lm(opioid_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)
```

```{r}

# filtering data by age to show effects of time on deaths in each age group

ages0_24 <- ageopioidmodel %>%
  filter(age_range == "0-24")

ages0_24 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages25_34 <- ageopioidmodel %>%
  filter(age_range == "25-34")

ages25_34 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages35_44 <- ageopioidmodel %>%
  filter(age_range == "35-44")

ages35_44 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages45_54 <- ageopioidmodel %>%
  filter(age_range == "45-54")

ages45_54 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages55_up <- ageopioidmodel %>%
  filter(age_range == "55+")

ages55_up %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

# looking at average deaths in each age group

averageageopioidbyage <- ageopioidmodel %>%
  drop_na() %>%
  group_by(age_range) %>%
  summarize(mean_deaths = mean(number_of_deaths))

# showing effects of being in each age group on number of deaths idt this makes sense

averageageopioid_model <- lm(number_of_deaths ~ age_range, data = ageopioidmodel)
averageageopioid_model %>%
  tidy(conf.int = TRUE) %>%
  select(term, estimate, conf.low, conf.high)
```