### 2020-05-07 Opioid_Crisis_Code.Rmd
moved data to a folder

---
title: "opioid_crisis_code"
author: "Suruchi Ramanujan"
date: "4/27/2020"
output: html_document
---

```{r setup, include=FALSE}

# I want to load all necessary packages.

library(dplyr)
library(janitor)
library(tidyverse)
library(ggthemes)
library(maps)
library(leaflet)
library(gganimate)
library(broom)
```

```{r commondiseases, echo = FALSE}

# total deaths per year in u.s. by cancer, drug overdose, cvd compared to 1999

deathsbycommondiseases <- read_csv("raw-data/deathsbycommondiseases.csv", 
                          col_types = "ldddddd") %>%
  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, 
         `Deaths by CVD`) %>%
  mutate(`Proportion of 1999 Deaths by Cancer` = 
           `Deaths by Cancer`/549829) %>%
  mutate(`Proportion of 1999 Deaths by Drugs` = 
           `Deaths by Drug Overdose`/19122) %>%
  mutate(`Proportion of 1999 Deaths by CVD` = 
           `Deaths by CVD`/949900)

# this is to look at the total levels of deaths based on disease; pivotted
# longer for ggplot

totaldeathsbycommondiseases <- deathsbycommondiseases %>%
  select(Year, `Deaths by Cancer`, `Deaths by Drug Overdose`, 
         `Deaths by CVD`) %>%
  pivot_longer(.,
             cols = starts_with("Deaths by"),
             names_prefix = "Deaths by",
             values_to = "Deaths") %>%
  rename(`Cause of Death` = name)

# just the info needed to map compared to 1999 proportions

propdeathsbycommondiseases <- deathsbycommondiseases %>%
  select(Year, `Proportion of 1999 Deaths by Cancer`, 
         `Proportion of 1999 Deaths by Drugs`, 
         `Proportion of 1999 Deaths by CVD`) %>%
  pivot_longer(.,
             cols = starts_with("Proportion of"),
             names_prefix = "Proportion of 1999 Deaths by",
             values_to = "Prop.Deaths") %>%
  rename(`Cause of Death` = name)
```

```{r ggplot commondiseases, echo = FALSE}

# plot year vs deaths for different diseases. transitions no longer work w
# plotly

plot_deathsbycommondiseases <- ggplot(totaldeathsbycommondiseases, aes(x = Year, 
                                        y = Deaths, color = `Cause of Death`)) +
  geom_line() +
  scale_color_viridis_d() +
  theme_classic() +
   labs(x = "Year",
       y = "Total Number of Deaths in the United States") +
  geom_point() +
  transition_reveal(Year)

# plot year vs proportion of 1999 deaths for different diseases. transitions no
# longer work w plotly

prop_deathsbycommondiseases <- ggplot(propdeathsbycommondiseases, 
                    aes(x = Year, y = Prop.Deaths, color = `Cause of Death`)) +
  geom_line() +
  scale_color_viridis_d() +
  theme_classic() +
  labs(x = "Year",
       y = "Proportion of 1999 Total Deaths in the United States") +
  geom_point() +
  transition_reveal(Year)

```

```{r overdose_counts, echo = FALSE}

# data from wonder from cdc..got rid of unreliable data

death_by_year_and_state <- read_csv("raw-data/Drug Overdose Deaths by Year and State.csv"
, col_types = "lcdccddddc") %>%
  filter(`Crude Rate` != "Unreliable")

# state info for map

us_state <- map_data("state") %>%
  rename(state_full = region)

# only want death by drugs not alcohol or unknown. join w state info

death_by_year_and_state <- death_by_year_and_state %>%
  filter(`UCD - Drug/Alcohol Induced Code` == "D") %>%
  rename(state_full = State) %>%
  mutate(state_full = tolower(state_full)) %>%
  left_join(us_state, by = "state_full" )

# change cruderate for plot

death_by_year_and_state$`Crude Rate` <-
  as.numeric(death_by_year_and_state$`Crude Rate`)

# make a map of opioid death rates by state

ggplot(data = death_by_year_and_state,
            mapping = aes(x = long, y = lat,
                          fill = `Crude Rate`, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A") +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

```

```{r notinuse, echo = FALSE}

#drug treatment centers

treatment_locations <- read_csv("raw-data/Drug Treatment Centers.csv", 
                                col_types = "ccccccccc") %>%
  select(`Program Name`, Street, City, State, Zipcode)

# state populations

state_pop <- read_csv("raw-data/State Population Estimate.csv", col_types = "dccdcdddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddddd") %>%
  rename(state_full = NAME) %>%
  mutate(state_full = tolower(state_full))

# https://simplemaps.com/data/us-cities longlatinfo

longlatinfo <- read_csv("raw-data/uscitieslonglat.csv", 
                        col_types = "ccccdcccddddcllcdcd") %>%
  rename(City = city) %>%
  rename(State = state_id)

# treatment location info

treatment_locations_map <- treatment_locations %>%
  left_join(longlatinfo, by = c("City", "State")) %>%
  group_by(State) %>%
  mutate(count = n()) %>%
  mutate(state_name = tolower(state_name)) %>%
  rename(state_full = state_name) %>%
  left_join(us_state, by = "state_full")

# modify data to be per capita

treatment_locations_map_per_cap <- treatment_locations_map %>%
  left_join(state_pop, by = "state_full") %>%
  group_by(State) %>%
  mutate(count_per_pop = n()/POPESTIMATE2019)

```

```{r treatmentcenters, echo = FALSE}

# map of distribution of treatment centers per capita

ggplot(data = treatment_locations_map_per_cap,
            mapping = aes(x = long, y = lat.y,
                          fill = count_per_pop, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,120)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom") +
  labs(color = "Number of Treatment Centers in State Per Capita")

ggsave("treatment_centers_per_capita.png")
```

```{r madeathrates, echo = FALSE}

# read in opioid related death rate per year of ma vs that of us

mavsus_death <- read_csv("raw-data/MAAgeAdjustedOpioidRelatedDeathRateByYear.csv", 
                         col_types = "cdc")

# only looking at data for which there is complete info

mavsus_death <- mavsus_death %>%
  group_by(Geography) %>%
  filter (Year != "2015") %>%
  filter (Year != "1999") %>%
  rename(deathperhundredthousand =
           `Age-Adjusted Opioid-Related Death Rate per 100,000 People`)

# ggplot grouped by geography

plot_mavsus_death <- ggplot(mavsus_death, 
      aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
  geom_line() +
  scale_color_viridis_d() +
theme_classic() +
  labs(x = "Year",
       y = "Deaths per 100,000",
       title = "Comparison of Death Rates between Massachusetts and the United States",
       subtitle = "Between the Years of 2000 and 2014") +
  geom_point()

```

```{r plot_1_ma, echo = FALSE}

# county data

counties <- read_csv("raw-data/zip_codes_states.csv",
                     col_types = "dddccc") %>%
  filter(state == "MA") %>%
  rename(County = county) %>%
  rename(Municipality = city)

# opioid related deaths by county

madeathbycounty <- read_csv(
  "raw-data/MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv",
                            col_types = "cdcdddc")

# county population for per capita info

countypop <- read_csv("raw-data/countypop.csv",
                      col_types = "cdd") %>%
  mutate(Pop = Pop/100000) %>%
  rename(subregion = CTYNAME) %>%
  mutate(subregion = tolower(subregion)) %>%
  select(subregion, Pop)

# counties + longitude/latitude

madeathbycountywlonglat <- madeathbycounty %>%
  left_join(counties, by = "Municipality")

# modifying and dividing by time and county/subregion

madeathbycountywlonglat <- madeathbycountywlonglat %>%
  select(County,
         `Confirmed Opioid Related Death Count 2001-2005`,
         `Confirmed Opioid Related Death Count 2006-2010`,
         `Confirmed Opioid Related Death Count 2011-2015`,
         `latitude`,
         `longitude`,
         `Municipality`) %>%
  distinct(Municipality, .keep_all= TRUE) %>%
  na.omit() %>%
  rename(subregion = County) %>%
  mutate(subregion = tolower(subregion)) %>%
  group_by(subregion) %>%
  mutate(total_deaths_2001.5 = 
          sum(`Confirmed Opioid Related Death Count 2001-2005`)) %>%
  mutate(total_deaths_2006.10 = 
          sum(`Confirmed Opioid Related Death Count 2006-2010`)) %>%
  mutate(total_deaths_2011.15 = 
           sum(`Confirmed Opioid Related Death Count 2011-2015`)) %>%
  distinct(subregion, .keep_all= TRUE)

# county map info

us_county <- map_data("county") %>%
  filter(region == "massachusetts")

# poverty by county (not in use as of now)

povertybycounty <- read_csv("raw-data/PovertyByCounty.csv",
                            col_types = "cd") %>%
  rename(subregion = County) %>%
  mutate(subregion = tolower(subregion))

# combining all mass data together

full_data <- us_county %>%
  left_join(madeathbycountywlonglat, by = "subregion") %>%
  left_join(countypop, by = "subregion") %>%
  left_join(povertybycounty, by = "subregion") %>%
  mutate(percap_2001.5 = total_deaths_2001.5/Pop) %>%
  mutate(percap_2006.10 = total_deaths_2006.10/Pop) %>%
  mutate(percap_2011.15 = total_deaths_2011.15/Pop)

# 2001-2005 poverty percent vs opioid death rate per capita by county

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2001.5)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# not in use but correlation between poverty rate and opioid death per capita

full_data %>%
  summarize(correlation = cor(percap_2001.5,`Poverty Percent`))

# 2006-2010 poverty percent vs opioid death rate per capita by county

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2006.10)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# not in use but correlation between poverty rate and opioid death per capita

full_data %>%
  summarize(correlation = cor(percap_2006.10,`Poverty Percent`))

# 2011-2015 poverty percent vs opioid death rate per capita by county

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2011.15)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# not in use but correlation between poverty rate and opioid death per capita

full_data %>%
  summarize(correlation = cor(percap_2011.15,`Poverty Percent`))

# poverty percent by county (not in use)

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = `Poverty Percent`,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,2,4,6,8,10)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

# ggplot of opioid death rate per capita by county between 01 and 05

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2001.5,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

# ggplot of opioid death rate per capita by county between 06 and 10

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2006.10,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

# ggplot of opioid death rate per capita by county between 11 and 15

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2011.15,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

# don't put these in final app bc they are the same graphs as above except w
# total death info rather than per capita- will prob confuse reader

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2001.5,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2006.10,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2011.15,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

```

```{r leaflet, echo = FALSE}

# just looking at treatment centers in MA

treatment_locations_map_mas <- treatment_locations %>%
  left_join(longlatinfo, by = c("City", "State")) %>%
  filter(State == "MA") %>%
  drop_na()

#treatment centers in MA distribution using leaflet

leaflet(options = leafletOptions(dragging = TRUE,
                                 minZoom = 8, 
                                 maxZoom = 10)) %>%
  addProviderTiles("CartoDB") %>%
  addCircleMarkers(data = treatment_locations_map_mas,
                   radius = 3,
                   label = ~`Program Name`)

```

```{r ageandrace, echo = FALSE}

# models generated from data from kff.org which took data from cdc

# age vs opioid deaths over time

ageopioidmodel <- read_csv("raw-data/U.S. age vs. opioid deaths.csv", 
                           col_types = "dcc") %>%
  clean_names() %>%
  filter(number_of_deaths != "N/A") %>%
  filter(age_range != "Total") %>%
  mutate(age_range = as.factor(age_range)) %>%
    mutate(number_of_deaths = as.numeric(number_of_deaths))

# race vs opioid deaths over time

raceopioidmodel <- read_csv("raw-data/U.S. race vs. opioid deaths.csv", 
                            col_types = "dcc") %>%
  clean_names() %>%
  na.omit() %>%
  mutate(race = as.factor(race)) %>%
  mutate(opioid_deaths = as.numeric(opioid_deaths))

# combine the two for ggplot later

ageandracemodel <- ageopioidmodel %>%
  na.omit() %>%
  full_join(raceopioidmodel, by = "year") %>%
  rename(deathsbyage = number_of_deaths) %>%
  rename(deathsbyrace = opioid_deaths)

# ggplot for age over time vs opioid deaths

ggplot(ageandracemodel, aes(year, deathsbyage, color = age_range)) +
  geom_point() +
  scale_color_viridis_d() +
  theme_classic() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Year", y = "Number of Opioid Deaths in America", 
       color = "Age Range")

# ggplot for race over time vs opioid deaths

ggplot(ageandracemodel, aes(year, deathsbyrace, color = race)) +
  geom_point() +
  scale_color_viridis_d() +
  theme_classic() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Race")
```

```{r race_lm, echo = FALSE}

# filtering data by race to show effects of time on deaths in each race group

death_vs_white <- raceopioidmodel %>%
  filter(race == "White, Non-Hispanic")

# all simple linear regression model

death_vs_white %>%
    lm(opioid_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high) %>%
    rename(Coefficient = estimate) %>%
    rename(`Lower End` = conf.low) %>%
    rename(`Upper End` = conf.high)

death_vs_black <- raceopioidmodel %>%
  filter(race == "Black, Non-Hispanic")

death_vs_black %>%
    lm(opioid_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

death_vs_hispanic <- raceopioidmodel %>%
  filter(race == "Hispanic")

death_vs_hispanic %>%
    lm(opioid_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)
```

```{r age_lm, echo = FALSE}

# filtering data by age to show effects of time on deaths in each age group

ages0_24 <- ageopioidmodel %>%
  filter(age_range == "0-24")

ages0_24 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages25_34 <- ageopioidmodel %>%
  filter(age_range == "25-34")

ages25_34 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages35_44 <- ageopioidmodel %>%
  filter(age_range == "35-44")

ages35_44 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages45_54 <- ageopioidmodel %>%
  filter(age_range == "45-54")

ages45_54 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages55_up <- ageopioidmodel %>%
  filter(age_range == "55+")

ages55_up %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

# looking at average deaths in each age group

averageageopioidbyage <- ageopioidmodel %>%
  drop_na() %>%
  group_by(age_range) %>%
  summarize(mean_deaths = mean(number_of_deaths))

# showing effects of being in each age group on number of deaths idt this makes
# sense so not including it yet

averageageopioid_model <- lm(number_of_deaths ~ age_range, 
                             data = ageopioidmodel)
averageageopioid_model %>%
  tidy(conf.int = TRUE) %>%
  select(term, estimate, conf.low, conf.high)
```